---
title: "Untitled"
author: "Eric Hirsch"
date: "11/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load libraries
library(tidymodels)
library(vip)
library(tidyverse)
library(lmtest)
library(skimr)
library(mltools)
library(psych)
```

```{r}
home_sales <- read.csv("D:\\RStudio\\605_Final\\Final\\housing\\train.csv", stringsAsFactors=TRUE, header = TRUE)
home_kaggle_test <- read.csv("D:\\RStudio\\605_Final\\Final\\housing\\test.csv", stringsAsFactors=TRUE, header = TRUE)
home_submit <- read.csv("D:\\RStudio\\605_Final\\Final\\housing\\sample_submission.csv", header = TRUE)
```
```{r}

head(home_sales, 10)
dim(home_sales)
glimpse(home_sales)
summary(home_sales)
skim(home_sales)
str(home_sales)
```


#### 1. Split the data

a. get rid of columns that are mostly na

```{r}

home_sales2 <- home_sales %>%
  dplyr::select_if(colSums(!is.na(.))>1200) 

```

remove utlities becasue it only has one value

```{r}

home_sales3 <- home_sales2 %>%
  dplyr::select(-Utilities) %>%
  dplyr::select(-Id)

```



#### 1. Split the data

Adjustments : 
1. log of y
2. eliminate na rows after eliminating mostly na columns
3. fix one-hots to account for sparse columns


```{r}

set.seed(271)

home_sales4 <- home_sales3 %>%
  mutate(SalePrice = log(SalePrice)) %>%
  na.omit(SalePrice)
```


Try again without tidymodels but using home_sales4 which already has log and na control

1. scale
2. one-hot
3. run regression
4. predict

```{r}
library(data.table)

home_sales10 <- one_hot(as.data.table(home_sales4), dropCols=TRUE)
home_sales11 <- as.data.frame(scale(home_sales10[-SalePrice]))

home_sales12 <- home_sales10 %>%
  dplyr::select_if(~ !any(is.na(.)))

describe(home_sales12$SalePrice)

```
```{r}

m1 = lm(SalePrice ~ ., data = home_sales12)
summary(m1)

par(mfrow=c(2,2)) # plot all 4 plots in one
plot(m1)

```
drop outliers

```{r}



```
predict
```{r}



home_kaggle_sales2 <- home_kaggle_test %>%
  dplyr::select_if(colSums(!is.na(.))>=1459) 

library(data.table)

home_kagglesales10 <- one_hot(as.data.table(home_kaggle_sales2), dropCols=TRUE)
home_kagglesales11 <- as.data.frame(scale(home_kagglesales10))

home_kagglesales12 <- home_kagglesales11 %>%
  dplyr::select_if(~ !any(is.na(.)))

#describe(home_kagglesales12)

```
```{r}
library(janitor)
cols <- janitor::compare_df_cols(home_sales12, home_kagglesales11)

cols1 <- cols %>%
  dplyr::filter(is.na(home_kagglesales11)) %>%
  dplyr::filter(column_name != "SalePrice")

cols2 <- as.vector(cols1$column_name)
cols2

home_sales13 <- home_sales12 %>%
  dplyr::select(-cols2)

```

```{r}
m2 = lm(SalePrice ~ ., data = home_sales13)
summary(m2)

par(mfrow=c(2,2)) # plot all 4 plots in one
plot(m2)
```
```{r}

prediction <- as.data.frame(predict(m2, newdata = home_kagglesales11))
prediction
prediction <- exp(prediction)
prediction



```

```{r}

library("writexl")
write_xlsx(prediction, "C:\\Users\\Eric\\Desktop\\people2.xlsx")
```

Deal with overfit

```{r}

df <- as.data.frame(summary(m2)$coefficients[])
df <- cbind(Coeffs = rownames(df), df)
rownames(df) <- 1:nrow(df)

df["p_value"] <- df[,5]

cols3 <- df %>%
  filter(p_value<.05) %>%
  arrange(desc(Estimate^2)) %>%
  dplyr::select(Coeffs) %>%
  filter(Coeffs != "(Intercept)") %>%
  head(200)

cols3[5,1] <- "RoofMatl_Tar&Grv"
#cols3 <-cols3[-7]
vCols1 <- cols3$Coeffs

vCols1 <- c("SalePrice", "OverallCond", "GrLivArea", "OverallQual", "RoofMatl_WdShngl", "Neighborhood_MeadowV", "Neighborhood_Edwards", "YearBuilt", "LotArea")

vCols1 <- c("SalePrice", "OverallCond", "GrLivArea", "OverallQual", "YearBuilt", "LotArea")

home_sales14 <- home_sales13 %>%
  dplyr::select(c(vCols1, SalePrice))

m4 = lm(SalePrice ~ ., data = home_sales14)
summary(m4)

par(mfrow=c(2,2)) # plot all 4 plots in one
plot(m4)

```

```{r}
prediction2 <- as.data.frame(predict(m4, newdata = home_kagglesales11))
prediction2
prediction2 <- exp(prediction2)
prediction2


library("writexl")
write_xlsx(prediction2, "C:\\Users\\Eric\\Desktop\\people.xlsx")
```

## Numeric
```{r}

numeric_data <- home_sales %>% select_if(is.numeric)
dim(numeric_data)
```
Keep most important

```{r}

dfAll <- numeric_data

numeric_data <- numeric_data[,c(1, 4, 7,8, 10:17, 28:34, 38)]
str(numeric_data)

```
plot to see

```{r}

numeric_data[, 2:20] %>%
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()


```
remove high zero counts

```{r}
df <- numeric_data[,-c(6,11,16:19)]
str(df)

```

Model
```{r}

# Standardize predictors
means <- sapply(df[,2:13],mean)
stdev <- sapply(df[,2:13],sd)
df.scaled <- as.data.frame(scale(df[,2:13], center=means, scale=stdev))
df.scaled$SalePrice <- df$SalePrice
df.scaled$Id <- df$Id
head(df.scaled)
```

fit

```{r}

attach(df.scaled)
mod_1 <- lm(SalePrice ~ LotArea + YearBuilt + YearRemodAdd + BsmtFinSF1 + BsmtUnfSF + TotalBsmtSF + X1stFlrSF + X2ndFlrSF + GrLivArea + GarageArea + WoodDeckSF + OpenPorchSF)
summary(mod_1)

```
Final model 
```{r}


mod_4 <- lm(SalePrice ~ LotArea + YearBuilt + YearRemodAdd + BsmtFinSF1 + TotalBsmtSF + X1stFlrSF + X2ndFlrSF + GarageArea + WoodDeckSF)
summary(mod_4)
```


Load test data

```{r}

#Load the data and remove columns same as our training data
test_df <- home_kaggle_test
test_df <- test_df %>% select_if(is.numeric)
test_df <- test_df[,c(1, 4, 7,8, 10:17, 28:34)]
test_df <- test_df[,-c(6,11,16:19)]
str(test_df)

```

Scale everything

```{r}

# Standardize test predictors
test.scaled <- as.data.frame(scale(test_df[,2:13], center=means, scale=stdev))
test.scaled$SalePrice <- test_df$SalePrice
test.scaled$Id <- test_df$Id
head(test.scaled)
```

Predict everything
```{r}

predictions <- predict(mod_4,newdata=test.scaled)
predictions <- data.frame(as.vector(predictions))
predictions$Id <- test.scaled$Id
predictions[,c(1,2)] <- predictions[,c(2,1)]
colnames(predictions) <- c("Id", "SalePrice")
predictions[is.na(predictions)] <- 0
head(predictions)
```

```{r}

library("writexl")
write_xlsx(predictions, "C:\\Users\\Eric\\Desktop\\people3.xlsx")
```
Try the log

```{r}
mod_5 <- lm(log(SalePrice) ~ LotArea + YearBuilt + YearRemodAdd + BsmtFinSF1 + TotalBsmtSF + X1stFlrSF + X2ndFlrSF + GarageArea + WoodDeckSF)
summary(mod_5)
```

Predict everything
```{r}
predictions <- predict(mod_5,newdata=test.scaled)
predictions <- data.frame(as.vector(predictions))
predictions$Id <- test.scaled$Id
predictions[,c(1,2)] <- predictions[,c(2,1)]
colnames(predictions) <- c("Id", "SalePrice")
predictions[is.na(predictions)] <- 0
predictions$SalePrice <- exp(predictions$SalePrice)
head(predictions)
```

```{r}
write.csv(predictions,"C:\\Users\\Eric\\Desktop\\people3.csv", row.names=FALSE)
```
Add neighborhood

```{r}

dfNeighborhood <- one_hot(as.data.table(home_sales), dropCols=TRUE)
dfNeighborhood <- dplyr::select(dfNeighborhood, Neighborhood_Blmngtn:Neighborhood_Veenker)

df6 <- dplyr::select(df.scaled, c(SalePrice, LotArea, YearBuilt, YearRemodAdd, BsmtFinSF1, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, GarageArea, WoodDeckSF))

dfNeighborhood_test <- one_hot(as.data.table(home_kaggle_test), dropCols=TRUE)
dfNeighborhood_test <- dplyr::select(dfNeighborhood_test, Neighborhood_Blmngtn:Neighborhood_Veenker)

test_df6 <- dplyr::select(test.scaled, c(LotArea, YearBuilt, YearRemodAdd, BsmtFinSF1, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, GarageArea, WoodDeckSF))

df6 <- cbind(df6, dfNeighborhood)
test_df6 <- cbind(test_df6, dfNeighborhood_test)

```

Try model with neighborhood

```{r}
mod_6 <- lm(log(SalePrice) ~ ., data=df6)
summary(mod_6)
```
predict 6
```{r}
predictions <- predict(mod_6,newdata=test_df6)
predictions <- data.frame(as.vector(predictions))
predictions$Id <- test.scaled$Id
predictions[,c(1,2)] <- predictions[,c(2,1)]
colnames(predictions) <- c("Id", "SalePrice")
predictions[is.na(predictions)] <- log(median(home_sales$SalePrice))
predictions$SalePrice <- exp(predictions$SalePrice)
head(predictions)
```
```{r}
write.csv(predictions,"C:\\Users\\Eric\\Desktop\\people6.csv", row.names=FALSE)
```

Add Roof

```{r}

dfRoof <- one_hot(as.data.table(home_sales), dropCols=TRUE)
dfRoof <- dplyr::select(dfRoof, c(RoofMatl_WdShngl, RoofMatl_CompShg, 'RoofMatl_Tar&Grv'))

dfRoof_test <- one_hot(as.data.table(home_kaggle_test), dropCols=TRUE)
dfRoof_test <- dplyr::select(dfRoof_test, c(RoofMatl_WdShngl, RoofMatl_CompShg, 'RoofMatl_Tar&Grv'))

df7 <- cbind(df6, dfRoof)
df7 <- dplyr::select(df7, -Neighborhood_Veenker)

test_df7 <- cbind(test_df6, dfRoof_test)
test_df7 <- dplyr::select(test_df7, -Neighborhood_Veenker)

```

Try model Roof material

```{r}
mod_7 <- lm(log(SalePrice) ~ ., data=df7)
summary(mod_7)
```
Predict 7
```{r}

predictions <- predict(mod_7,newdata=test_df7)
predictions <- data.frame(as.vector(predictions))
predictions$Id <- test.scaled$Id
predictions[,c(1,2)] <- predictions[,c(2,1)]
colnames(predictions) <- c("Id", "SalePrice")
predictions[is.na(predictions)] <- log(median(home_sales$SalePrice))
predictions$SalePrice <- exp(predictions$SalePrice)
head(predictions)
```

```{r}

write.csv(predictions,"C:\\Users\\Eric\\Desktop\\people7.csv", row.names=FALSE)
```

```{r}
plot(mod_7)
```

`Remove outliers

```{r}
df8 <- slice(df7, -1299)
df8 <- slice(df8, -524)
```

```{r}

mod_8 <- lm(log(SalePrice) ~ ., data=df8)
summary(mod_8)
```

predict 8

```{r}

predictions <- predict(mod_8,newdata=test_df7)
predictions <- data.frame(as.vector(predictions))
predictions$Id <- test.scaled$Id
predictions[,c(1,2)] <- predictions[,c(2,1)]
colnames(predictions) <- c("Id", "SalePrice")
predictions[is.na(predictions)] <- log(median(home_sales$SalePrice))
predictions$SalePrice <- exp(predictions$SalePrice)
head(predictions)
```

```{r}

write.csv(predictions,"C:\\Users\\Eric\\Desktop\\people8.csv", row.names=FALSE)
```

```{r}
library(MASS)
step_lm <- stepAIC(mod_8, trace=FALSE)
summary(step_lm)

```

```{r}
dfAll2 <- na.omit(dfAll)

mod_9 <- lm(log(SalePrice) ~ ., data=dfAll2)
summary(mod_9)
```
```{r}
step_lm2 <- stepAIC(mod_9, trace=FALSE)
summary(step_lm2)

```

Add neighborhood 2

```{r}

dfAll = scale(dfAll)

dfNeighborhood <- one_hot(as.data.table(home_sales), dropCols=TRUE)
dfNeighborhood <- dplyr::select(dfNeighborhood, Neighborhood_Blmngtn:Neighborhood_Veenker)

df6 <- dfAll

dfNeighborhood_test <- one_hot(as.data.table(home_kaggle_test), dropCols=TRUE)
dfNeighborhood_test <- dplyr::select(dfNeighborhood_test, Neighborhood_Blmngtn:Neighborhood_Veenker)

test_df6 <- home_kaggle_test %>% select_if(is.numeric)

df6 <- cbind(df6, dfNeighborhood)
test_df6 <- cbind(test_df6, dfNeighborhood_test)



dfRoof <- one_hot(as.data.table(home_sales), dropCols=TRUE)
dfRoof <- dplyr::select(dfRoof, c(RoofMatl_WdShngl, RoofMatl_CompShg, 'RoofMatl_Tar&Grv'))

dfRoof_test <- one_hot(as.data.table(home_kaggle_test), dropCols=TRUE)
dfRoof_test <- dplyr::select(dfRoof_test, c(RoofMatl_WdShngl, RoofMatl_CompShg, 'RoofMatl_Tar&Grv'))

df16 <- df6

df17 <- cbind(df16, dfRoof)
df17 <- dplyr::select(df17, -Neighborhood_Veenker)

test_df17 <- cbind(test_df6, dfRoof_test)
test_df17 <- dplyr::select(test_df17, -Neighborhood_Veenker)

df17 <- na.omit(df17)


```

Try model with neighborhood

```{r}

df17 <- slice(df17, -1299)
df17 <- slice(df17, -524)

mod_17 <- lm(log(SalePrice) ~ ., data=df17)
summary(mod_17)
```


```{r}
step_lm <- stepAIC(mod_17, trace=FALSE)
summary(step_lm)

```

predict 17
```{r}
predictions <- predict(mod_17,newdata=test_df17)
predictions <- data.frame(as.vector(predictions))
predictions$Id <- test.scaled$Id
predictions[,c(1,2)] <- predictions[,c(2,1)]
colnames(predictions) <- c("Id", "SalePrice")
predictions[is.na(predictions)] <- log(median(home_sales$SalePrice))
predictions$SalePrice <- exp(predictions$SalePrice)
head(predictions)
```

```{r}

plot(step_lm)
```

```{r}
write.csv(predictions,"C:\\Users\\Eric\\Desktop\\people17.csv", row.names=FALSE)
```
Use algorithm to find most important
```{r}
#get numeric_data and scale
x1 <- as.data.frame(scale(dfAll[,1:37]))

x1 = na.omit(x1)

#Run on log
mx1 <- lm(log(SalePrice) ~ ., data=x1)

#4 step

step_lm <- stepAIC(mx1, trace=FALSE)
summary(step_lm)

#5 gather most important
```

