---
title: "Eric_Hirsch_605_Assignment_13"
author: "Eric Hirsch"
date: "11/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. __*Use integration by substitution to solve the integral below.*__

$$
\int { 4{ e }^{ -7x }dx } 
$$

We know that:
$$
\int {e^xdx = e^x + C}
$$

If we set u to -7x, then du = -7dx

So now we have: 
$$
\int { 4{ e }^{ -7x }dx }\\
4 \int { { e }^{ -7x }dx } \\
\frac{-4}{7} \int { { e }^{ u }du } \\
\frac{-4}{7}e^{u}+c\\
\frac{-4}{7}e^{-7x}+c
$$



2. __*Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of*__

$$
\frac{dN}{dt}=\frac{-3150}{t^{4}}-220
$$
__*bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N( t ) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.*__

This is simply a matter of integrating the rate function.  The level after one day will allow us to find the constant.

$$
 \int {(-3150{t^{-4}}-220)dt}\\ = 
  -1/3(-3150t)^{-3}-220t+c\\ =
{1050}{t^{-3}}-220t+c
$$

Now we solve for c:

$$
\frac{1050}{1^{3}}-220(1)+c=6350\\
c = 6530-1050+220\\
c=5700
$$

So here is the function:

$$
N(t)=  \frac{1050}{t^{3}}-220t+5700
$$



 
3. Find the total area of the red rectangles in the figure below, where the equation of the line is f ( x ) = 2x - 9.

We can do this a numbeer of ways, but since it's a calculus assignment we'll integrate. We assume the interval is 4.5 to 8.5 from the graph (although this isn't 100% certain):


$$
\int _{ 4.5 }^{ 8.5} \ 2x-9 \ { dx }\\
$$
Integral:
$$x^2 - 9x$$
Thus:
$$(8.5^2 - 9(8.5)) - (4.5^2-9(4.5)) = 16$$
4. __*Find the area of the region bounded by the graphs of the given equations.*__

$$
y=x^{2}-2x-2\\
y=x+2
$$
For this question first we find the points of intersection:
$$
x^{2}-2x-2=x+2\\
x^{2}-3x-4=0\\
(x-4)(x+1)=0\\
$$
x=4 and x=-1.

So we need to subtract the area under the upside down parabola from all the area under the line - this will give us the area within the parabola.

There are many ways to get the area under the line - we could do 1*5 (the rectangular part of the area) plus 25/2 (The triangular part of the area.)  This gives us 17.5.

Now we subtract:

```{r}
int2 <- function(x) {(x^2-2*x-2)}
integrate(int2, lower = -1, upper = 4)
```
17.5 - -3.33333 = 20.8333333

__*5. A beauty supply store expects to sell 110 flat irons during the next year. It costs 3.75 to store one flat iron for one year. There is a fixed cost of 8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.*__

I think we need to make some assumptions here about rates of purchase. Setting n = # of orders, we will assume that storage costs are (3.75*110)/n - in other words, items bought for sale remain in the store until the next order and are bought on the day of the new order.

Thus, 

$$
T=8.25n + \frac{3.75*110}{n}\\
$$
This is a minimization problem, so we take the first derivative and set to 0:

$$
T'=8.25 + -1(\frac{3.75*110}{n^2}) = 0\\
8.25*x{^2} = (3.75*110)\\
x{^2} = (3.75*110)/8.25\\
x = sqrt((3.75*110)/8.25) = ~ 7
$$
This gives us around 7 lots, with around (110/7 = 16) irons per lot.

6. __*Use integration by parts to solve the integral below*__

$$
\int { ln(9x)x^{6}dx } 
$$

To integrate by parts we use this formula:

$$
uv-\int vdu \\
$$
We will set u = ln(9x) and we can set dv = x^6.  Now we need du and v:

$$
u=ln(9x)\\
du=\frac{1}{x}dx\\
$$

Now we calculate v by integrating dv:
$$
dv=x^{6}\\
v=\frac{1}{7}x^{7}
$$

This gives us:

$$
\frac{x^{7}ln(9x)}{7}-\int \frac{1}{7}x^{7}\frac{1}{x}dx\\
$$
So now we simplify.  To make the notation easier I'm going to set the first element to q:

$$
q = \frac{x^{7}ln(9x)}{7}
$$
So now we have:
$$
q-\int \frac{x^6}{7}dx\\
q-\frac{1}{7}\int x^{6}dx\\
q-\frac{1}{7}(\frac{x^{7}}{7})+C\\
q-\frac{x^{7}}{49}+C\\
\frac{x^{7}ln(9x)}{7}-\frac{x^{7}}{49}+C\\
\frac{1}{7}({x^{7}ln(9x)}-\frac{x^{7}}{7})+C
$$

__*7.   Determine whether f ( x ) is a probability density function on the interval 1, e^6*__

$$
f(x)=\frac{1}{6x}
$$

A probability density function will have an integral of 1 and be positive for all possible values - so that is what we are testing here.  Positivity is guaranteed because the interval is positive and there are no negative signs in the function.  So now we integrate:



$$
\int _{ 1 }^{ { e }^{ 6 } }{ \frac { 1 }{ 6x } dx } \\
\frac{1}{6} \int _{ 1 }^{ { e }^{ 6 } }{ \frac { 1 }{ x } dx } \\
$$
Here we use R to generate the integral:

```{r}
int2 <- function(x) {1/x}
integrate(int2, lower = 1, upper = exp(1)^6)
```

Because 1/6*6 = 1, then the distribution can be regarded as a probability density function.
