---
title: "Assignment 3"
author: "Eric Hirsch"
date: "7/20/2021"
output: html_document
---

# Problemset 1

### 1. What is the rank of the matrix A?
 
```{r}
#Use the qr function to determine the rank
A <- matrix(data = c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow = 4, byrow=T)
A

qr(A)$rank

#We can also confirm that the determinant is non-zero - thus the rank is the number of rows (or columns) for a square matrix, which is 4.
det(A)
```

### 2. Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?
 
The maximum rank of a matrix is equal to the lower of the number of rows and the number of columns - since n is lower than m, n is the maximum rank.  The minimum rank for a nonzero matrix is 1.

### 3. What is the rank of matrix B?

```{r}
B <- matrix(c(1,2,1,3,6,3,2,4,2), nrow = 3, byrow=T)
B
qr(B)$rank
```

The rank is 1.  We can easily see the linear dependence - row 2 is simply triple row 1 and row 3 is double row 1. 

# Problemset 2

### 1. Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

### a. Compute the Eigenvalues

$$ A = \begin{bmatrix}1 & 2 & 3 \\0 & 4 & 5 \\0 & 0 & 6 \end{bmatrix}$$ 

So in other words, using the characteristic polynomial, we need to solve for lambda where ${det}\left(A-\lambda \,I\right)\,=\,0$

${det}\left(\begin{bmatrix}1&2&3\\0&4&5\\0&0&6\end{bmatrix} - \begin{bmatrix}\lambda&0&0\\0&\lambda&0\\0&0&\lambda\end{bmatrix}\right)=0$

Thus, 

${det}\left(\begin{bmatrix}1-\lambda &2&3\\0&4-\lambda &5\\0&0&6-\lambda \end{bmatrix}\right)=0$

Because matrix A is an upper triangle matrix the formula for the determinant is greatly simplified since most of the terms will zero out. In fact, the determinant of an upper triangular matrix is the product of its diagonal elements ^[1]. What is left, then, is $\left(1-\lambda\right)\left(4-\lambda \right)\left(6-\lambda \right)$

Setting this to 0, we can see that 1, 4 or 6 would need to be 0 - thus the eigenvalues are 1, 4 and 6.

### b. Compute the Eigenvectors for Each Eigenvalue

#### Eigenvalue = 1

$\begin{bmatrix}1-\lambda &2&3\\0&4-\lambda &5\\0&0&6-\lambda \end{bmatrix}\begin{bmatrix}X\\Y\\Z\end{bmatrix}=\begin{bmatrix}0&2&3\\0&3&5\\0&0&5\end{bmatrix}\begin{bmatrix}X\\Y\\Z\end{bmatrix}=0$

Thus:

    2y + 3z=0
    3y + 5z=0
    5z = 0
    
    If we set x=1 and both y and z must be zero, the eigenvector is:
    
$\begin{bmatrix}1\\0\\0\end{bmatrix}$


#### Eigenvalue = 4

$\begin{bmatrix}1-\lambda &2&3\\0&4-\lambda &5\\0&0&6-\lambda \end{bmatrix}\begin{bmatrix}X\\Y\\Z\end{bmatrix}=\begin{bmatrix}-3&2&3\\0&0&5\\0&0&2\end{bmatrix}\begin{bmatrix}X\\Y\\Z\end{bmatrix}=0$

Thus:

    -3x + 2y + 3z=0
    5z=0
    2z = 0
    
    If z = 0 and we set y=1, then -3x + 2 = 0, x = 2/3
    
    The eigenvector is:
    
$\begin{bmatrix}2/3\\1\\0\end{bmatrix}$

#### Eigenvalue = 6

$\begin{bmatrix}1-\lambda &2&3\\0&4-\lambda &5\\0&0&6-\lambda \end{bmatrix}\begin{bmatrix}X\\Y\\Z\end{bmatrix}=\begin{bmatrix}-5&2&3\\0&-2&5\\0&0&0\end{bmatrix}\begin{bmatrix}X\\Y\\Z\end{bmatrix}=0$

Thus:

    -5x + 2y + 3z=0
    -2y + 5z=0
    
    If we set z=1, 
    then -2y + 5 = 0, y = 5/2
    then -5x +5 +3 = 0, -5x + 8 = 0, x=8/5
    
    
    The eigenvector is:
    
$\begin{bmatrix}8/5\\5/2\\1\end{bmatrix}$

1 This Khan Academy video (https://www.khanacademy.org/math/linear-algebra/matrix-transformations/determinant-depth/v/linear-algebra-upper-triangular-determinant) provided a good explanation of this for me.